---
title: "Project Two"
output:
  pdf_document: default
  html_document: default
---

Due Oct. 13 at 11:59 PM. 

For this first part of the exam, you can either use `surveys_complete.csv` or your own data. If you are using your own data, you must have data in which you think you have a numerical predictor variable and a numerical response variable. If you are using `surveys_complete`, you can use weight and hindfoot_length for this.

Save this file in your `projects` directory. You can either save it in a project two subdirectory, or just put it in projects. Either way is fine.


1) Load in your data. Which variable will you be using as a predictor, and which as a response? (5 pts)

```{r}
sodium_data <- read_csv("../../sodium_data/SodiumChannelData2.csv")
head(sodium_data)
```

```
Slow_dev_t1_ms is the predictor and SSSI_V0.5_mV is response
```

2) Plot the two against each other with a scatter plot. Do the data appear to be related linearly? (5 pts)


```{r}
ggplot(sodium_data, aes(x = Slow_dev_t1_ms, y = SSSI_V0.5_mV)) + geom_point()
```

```
I do not think that the data is related linearly. 
```


3) Fit the linear model. View the summary. (5 pts)


```{r}
sodium_line_fit <- lm(SSSI_V0.5_mV ~ Slow_dev_t1_ms, data = sodium_data)
summary(sodium_line_fit)
```

4) Does the summary make sense when you compare it to the plot you made? Does our model have good predictive power? Evaluate the residual standard error, intercept, and R-Squared in particular. Would you say your predictor predicts the response?  (10 pts)


```
The summary does make sense for the plot, and the summary and plot both demonstrate that the model does not have good predicting power. The residual standard error is pretty large for the sample, which makes sense because many of the points are not very close to the line. The intercept makes sense for the V 1/2 measures of the response. The R-squared is average if not a bit low, so with that and the insignificant p-value, I would be confident in the conclusion that there is not a linear relationship.
```


5) Plot the model on the graph. Increase the size of the text so it is comfortably readable at 5 feet. Make sure axis labels are readable and not overlapping with one another. (5 pts)

```{r}
sodium_plot <- ggplot(sodium_data, aes(x = Slow_dev_t1_ms, y = SSSI_V0.5_mV)) + geom_point(size = 0.5)
sodium_plot + geom_smooth(method = "lm", color = "blue", linewidth = 0.5, fill = "deeppink") + annotate("text", x = 700, y = 30, label = "R^2 == 0.431", parse = T) + theme_bw()
```


6) Check the normality of the residuals. Do they look OK? Are we violating assumptions? (5 pts)

```{r}
sodium_line_fit <- lm(SSSI_V0.5_mV ~ Slow_dev_t1_ms, data = sodium_data)
augmented_sodium <- augment(sodium_line_fit)
qqnorm(augmented_sodium$.resid)
qqline(augmented_sodium$.resid)

# These data are not normally distributed, and they are representative of the fact that these data do not fit a linear model well.
```

Why is normality of residuals important? 

```
By ensuring that the residuals are normally distributed, we can be sure that the model fits the majority of the data well. 
```

7) If you are using `surveys_complete`: Is there interspecific variation in the linear model? Hint: look at our prior work with facet plots. (15 pts) 

If you are *not* using  `surveys_complete`: Do you think there are groupings within your data that may have a separate linear model? Try grouping the data by that variable and redoing the lm. If this would not make sense for your data, you may also attempt to do multiple predictor variables. (15 pts)

```{r}
sodium_data %>% 
  group_by(Channel, Effects_Sact, SSFI_V0.5_mV) %>% 
  ggplot(aes(x = Channel, y = SSFI_V0.5_mV, color = Effects_Sact)) + geom_point()
```

## Part Two

In this portion, you are free to use your own data if you have a categorical predictor variable and a response variable. Otherwise, you may use the columns sex and weight in `surveys_complete`

1) First, plot the data grouped by sex (5 pts)

```{r}
ggplot(sodium_data, aes(x = Effects_Sact, y = SSSI_V0.5_mV, color = Effects_Sact)) + geom_jitter()
```

2) Try an ANOVA of this data (5 pt)

```{r}
anova_sodium_fit <- aov(sodium_line_fit1)
summary(anova_sodium_fit)
```

3) How about a linear model? What information does this give you that an ANOVA can't? (5 pts)


```{r}
sodium_line_fit1 <- lm(SSSI_V0.5_mV ~ Effects_Sact, data = sodium_data)
summary(sodium_line_fit1)
```

```
The linear model provides more information about the significance of the relationships between the variables whereas the anova provides a broad picture of the relationship in the data.
```

3) Plot the lm with the data, with points colored by sex. (10 pts)


```{r}
sodium_plot1 <- ggplot(sodium_data, aes(x = Effects_Sact, y = SSSI_V0.5_mV, color = Effects_Sact)) + geom_point()
sodium_plot1 + geom_smooth(method = "lm", color = "blue", linewidth = 0.5, fill = "deeppink")
```

4) Choose any model we've looked at so far, but use two predictor variables. Perform an LM, plot the results, and state if this changes your interpretation of the relationship between variables (10 pts)

```{r}
sodium_line_fit2 <- lm(SSSI_V0.5_mV ~ Slow_dev_t1_ms * Effects_Sact, data = sodium_data)
summary(sodium_line_fit2)
```

```{r}
sodium_plot2 <- ggplot(sodium_data, aes(x = Slow_dev_t1_ms, y = SSSI_V0.5_mV, color = Effects_Sact)) + geom_point()
sodium_plot2 + geom_smooth(method = "lm", color = "yellow", linewidth = 0.5, fill = "purple")
```

```
Due to the lack of individual data points, this plot does not change anything about what we can already see in the data.
```

## Part Three


1) Add and commit this document (5 pts)

```
After clicking the "Git" tab next to the Environment, I clicked the "Commit" button (white paper with a green check mark).
```

2) Push your changes to github (10 pts)

```
When you are ready to push, just hit the green arrow on the Git tab.
```



# MS students

My expectation is that you'll do this with your own data. If any part of this doesn't make sense with your data, please get in touch ASAP so we can work it out.
